Tool-driven Revolutions in Archaeological Science

### 2018-09-09 analysis/paper/jcaa-paper.Rmd
added to discussion what kind of revolution we are talking about

The use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R. To make it easier for archaeologists and researchers in other fields to use R for research and publication, we describe the R package rrtools ('reproducible research tools'). This package aims to simplify many of the steps needed to write reproducible research, and to guide users to best practices with minimal effort. The rrtools package contains functions that create a file structure according to fundamental principles of organising files for research. First, we keep the data separate from the methods. This means keep the code in a separate directory to the data. Second, we, keep the raw data seprate from data that is created or derived during the analysis.These two basic principles make it easy to stay organised during a complicated project, and make it easier for a reader to navigate their way through the compendium of files that are shared with the publication.

>>>>

The use of R is on the rise among archaeologists using natural and physical science methods, and has concrete advantages for authors that publish papers that cite R as they tend to be cited more often. This also is an indicator for the increase in popularity of R.

The use of R or any other scripting languages enables a paradigm-shift, as it offers the possibility to share one's code. The sharing of code has been identified in other disciplines as an amplifier of methodological change [@touchon_and_mccoy_2016]. It also enables researchers to not just reproduce the work of others, but to also evaluate them according to the scientific method. Thus, the sharing of code may be at the root of a tool-driven paradigm-change as desribed in the work of @Galison_1997.

Speaking with Kuhn's view on paradigm-change, a change in the view point of researchers is necessary. We propose that a change of point of view of the archaeological researcher is needed: She or he, who works with computational methods, needs to view him/herself as a scientist, who needs to adhere to the scientific method, i. e. making the research reproducible for others.

We acknowledge, this is not an easy task. Learning a scripting language is time-intensive and writing code so that one is comfortable in sharing it even more so. [some more info from your case study maybe?]

To make it easier for archaeologists and researchers in other fields to use R for research and publication, we describe the R package rrtools ('reproducible research tools'). This package aims to simplify many of the steps needed to write reproducible research, and to guide users to best practices with minimal effort. The rrtools package contains functions that create a file structure according to fundamental principles of organising files for research. First, we keep the data separate from the methods. This means keep the code in a separate directory to the data. Second, we, keep the raw data seprate from data that is created or derived during the analysis.These two basic principles make it easy to stay organised during a complicated project, and make it easier for a reader to navigate their way through the compendium of files that are shared with the publication.
